\bf Model Visualization Techniques for a Social Network Model

### 2018-08-15 paper-child.Rnw
fix a fig ref

set.seed(12345)
ggplot(data = jTT, aes(from_id = from, to_id = to)) +
  geom_net(aes(shape = as.factor(group)), directed = T, labelon = T,
           labelcolour='white',vjust = 0.5, hjust =0.5, arrowgap = .15,
           size=10, fontsize = 5, ecolour="black",
           linetype = c("solid", "solid","dotted", "solid")) +
  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
  ThemeNet +
  theme(legend.position = "none") + 
  labs(title = expression(paste(s[i3], "(x,z)", "=",Sigma[j != h], x[ij], x[ih], x[hj] %.% 
                                I, "(", z[i], "=", z[h]!= z[j], ")")
                          )
       )
@
\caption{\label{fig:jtt}Realization of a JTT, where $i$ is the ego, $j$ is the alter, and $h$ is the intermediary. The covariate is represented by the shape of the nodes.}
\end{subfigure}\hfill
\begin{subfigure}[t]{.45\textwidth}
<<dab, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center', out.width='.49\\textwidth'>>=
dade <- data.frame(from = c('i', 'i', 'h', 'k', 'i'), to = c('h', 'k', 'j', 'j','j'))
dadn <- data.frame(id = letters[8:11], group = c(1,1,1,1))

@@

To reorder the vertices for the matrix visualization, we first constructed a cumulative adjaceny matrix, $\mathbf{A}^{cum}$, for the series of microsteps simulating the network from $x(t_m)$ to $x(t_{m+1})$. A single entry in the cumulative adjacency matrix, $\mathbf{A}^{cum}_{ij}$, is the total number of times the edge $i \rightarrow j$ appears in the network from the initial observation, $x(t_m) \equiv X(0)$ to the final result of the last microstep, $X(R)$, where $R$ is the total number of microsteps taken:
\begin{equation}\label{eq:acum}
\mathbf{A}^{cum}_{ij} = \sum_{r = 0}^R X_{ij}(r). 
\end{equation}
We then performed a principal component analysis (PCA) on $\mathbf{A}^{cum}$, and used the values of the first principal component to order the vertices on the $x$ and $y$ axes for the adjacency matrix animation. In Figure~\ref{fog:adjmatorder}, we present the first wave of data ordered by the arbitrary node ID order alongside the seriated adjacency matrix ordered by the first principal component loading on the cumulative adjacency matrix, $\mathbf{A}^{cum}$. Using PCA on $\mathbf{A}^{cum}$ to order the rows and columns of the adjacency matrix visualization clearly shows the two distinct connected components in the first wave of the network, which are difficult to find in the arbitrarily ordered visualization.

>>>>

set.seed(12345)
ggplot(data = jTT, aes(from_id = from, to_id = to)) +
  geom_net(aes(shape = as.factor(group)), directed = T, labelon = T,
           labelcolour='white',vjust = 0.5, hjust =0.5, arrowgap = .15,
           size=10, fontsize = 5, ecolour="black",
           linetype = c("solid", "solid","dotted", "solid")) +
  expand_limits(x=c(-0.1,1.1), y=c(-0.1,1.1)) +
  ThemeNet +
  theme(legend.position = "none", plot.title = element_text(size = rel(0.9))) + 
  labs(title = expression(paste(s[i3], "(x,z)", "=",Sigma[j != h], x[ij], x[ih], x[hj] %*% 
                                I, "(", z[i], "=", z[h]!= z[j], ")")
                          )
       )
@
\caption{\label{fig:jtt}Realization of a JTT, where $i$ is the ego, $j$ is the alter, and $h$ is the intermediary. The covariate is represented by the shape of the nodes.}
\end{subfigure}\hfill
\begin{subfigure}[t]{.45\textwidth}
<<dab, echo=FALSE, fig.width=2, fig.height=2, message = FALSE, warning = FALSE, fig.align='center', out.width='.49\\textwidth'>>=
dade <- data.frame(from = c('i', 'i', 'h', 'k', 'i'), to = c('h', 'k', 'j', 'j','j'))
dadn <- data.frame(id = letters[8:11], group = c(1,1,1,1))

@@

To reorder the vertices for the matrix visualization, we first constructed a cumulative adjaceny matrix, $\mathbf{A}^{cum}$, for the series of microsteps simulating the network from $x(t_m)$ to $x(t_{m+1})$. A single entry in the cumulative adjacency matrix, $\mathbf{A}^{cum}_{ij}$, is the total number of times the edge $i \rightarrow j$ appears in the network from the initial observation, $x(t_m) \equiv X(0)$ to the final result of the last microstep, $X(R)$, where $R$ is the total number of microsteps taken:
\begin{equation}\label{eq:acum}
\mathbf{A}^{cum}_{ij} = \sum_{r = 0}^R X_{ij}(r). 
\end{equation}
We then performed a principal component analysis (PCA) on $\mathbf{A}^{cum}$, and used the values of the first principal component to order the vertices on the $x$ and $y$ axes for the adjacency matrix animation. In Figure~\ref{fig:adjmatorder}, we present the first wave of data ordered by the arbitrary node ID order alongside the seriated adjacency matrix ordered by the first principal component loading on the cumulative adjacency matrix, $\mathbf{A}^{cum}$. Using PCA on $\mathbf{A}^{cum}$ to order the rows and columns of the adjacency matrix visualization clearly shows the two distinct connected components in the first wave of the network, which are difficult to find in the arbitrarily ordered visualization.
